\subsection{Architecture Inspection Guide}\label{appendix:archInspection}
This list is a guide for reviewers in the peer review of the Module Guide (MG)
document (Section~\ref{sec:archVV}). The list organizes prompts by MG section
in the order that they appear in the document. Reviewers are encouraged to
include their own checks to this guide.

\paragraph{Introduction}
\begin{itemize}

    \item Does the introduction summarize the problem domain described in the
    SRS?

    \item Is the document's purpose described unambiguously?

    \item Is the document's purpose described accurately?

    \item Are the characteristics of the intended reader describing the readers
    of the MG?

    \item Are the characteristics of the intended reader complete?

    \item Is the document accessible to its intended readers?

    \item Is the description of the MG's document structure accurate?

\end{itemize}

\paragraph{Anticipated and Unlikely Changes}
\begin{itemize}

    \item Are the likely changes from the SRS associated with at least one
    anticipated change?

    \item Are the unlikely changes justified?

\end{itemize}

\paragraph{Connection Between Requirements and Design}
\begin{itemize}

    \item Is the justification for the proposed architecture unambiguous?

    \item Is the justification for the proposed architecture traceable to
    stakeholder needs as described in the SRS?

    \item Is the justification for the proposed architecture traceable to one
    or more functional and/or nonfunctional requirements from the SRS?

\end{itemize}

\paragraph{Module Hierarchy and Decomposition}
\begin{itemize}

    \item Do the modules satisfy the principle of information hiding?

    \item Is there one secret per module? If not, is it justified?

    \item Are the behaviour-hiding modules related to requirements in the SRS?

    \item Are the software decision modules related to concepts that are not in
    the SRS?

    \item Are the software decision modules necessary to the design?

\end{itemize}

\paragraph{Traceability Matrices}
\begin{itemize}

    \item Is every functional requirement from the SRS satisfied by at least
    one module?

    \item Is every module used to satisfy at least one functional requirement
    from the SRS?

    \item Is there one anticipated change per module? If not, is it justified?

\end{itemize}

\paragraph{Use Hierarchy Between Modules}
\begin{itemize}

    \item Is the use hierarchy complete?

    \item Is the use hierarchy accurate?

    \item Is the use hierarchy hierarchical?

    \item Does the hierarchy satisfy the principle of low coupling between
    modules?

\end{itemize}

\paragraph{Readability}
\begin{itemize}

    \item Is the document well structured?

    \item Is the writing clear and cohesive?

    \item Is the writing concise?

    \item Do all symbols, abbreviations, and notation in the MG appear in the
    Reference Material?

    \item Is the document free of spelling and grammar errors?

\end{itemize}