\subsection{Design Inspection Guide}\label{appendix:designInspection}
This list is a guide for reviewers in the peer review of the Module Interface
Specification (MIS) document (Section~\ref{sec:designVV}). The list organizes
prompts by MIS section in the order that they appear in the document. Reviewers
are encouraged to include their own checks to this guide.

\paragraph{Introduction}
\begin{itemize}

    \item Does the introduction summarize the problem domain described in the
    SRS?

    \item Does the introduction have references/links to the SRS and MG?

\end{itemize}

\paragraph{Notation}
\begin{itemize}

    \item Is the notation unambiguous?

    \item Is the notation complete?

    \item Is the notation used consistently throughout the MIS?

\end{itemize}

\paragraph{Module Hierarchy}
\begin{itemize}

    \item Does the module hierarchy match the MG?

    \item Is there a specification for each module in the module hierarchy?

\end{itemize}

\paragraph{Module Specifications}
\begin{itemize}

    \item Does the introductory content describe the purpose and key
    characteristics of the module?

    \item Are the modules classified correctly (e.g. Interface, ADT, Generic)?

    \item Are all exported constants defined with a literal value?

    \item Are the assumptions justified?

    \item Do all access routines do something (i.e. output or state transition)?

    \item Are all operations type safe?

    \item Are all state invariants initially satisfied?

    \item Are all state invariants satisfied when an access routine terminates?

    \item Do local functions make the specification easier to understand?

    \item Does each module satisfy the principle of high cohesion?

\end{itemize}

\paragraph{Readability}
\begin{itemize}

    \item Is the document well structured?

    \item Is the writing clear and cohesive?

    \item Is the writing concise?

    \item Do all symbols, abbreviations, and notation in the MIS appear/have a
    reference to a definition in the Reference Material?

    \item Is the document free of spelling and grammar errors?

\end{itemize}