\subsection{Master Test Plan Inspection Guide}\label{appendix:mtpInspection}
This list is a guide for reviewers in the peer review of the Master Test Plan
(MTP) document (Section~\ref{sec:mtpVV}). The list organizes prompts by MTP
section in the order that they appear in the document. Reviewers are encouraged
to include their own checks to this guide.

\paragraph{Introduction}
\begin{itemize}

    \item Does the introduction describe the purpose of \progname{}'s MTP?

    \item Does the summary of \progname{} describe its intended purpose?

    \item Does the summary of \progname{} list all of its required
    characteristics?

    \item Are \progname{}'s required characteristics consistent with its
    intended purpose?

    \item Are the aspects of \progname{} and its software development artifacts
    (SDA) that are to be tested unambiguous?

    \item Are the aspects of \progname{} and its SDAs that are \textit{not} to
    be tested unambiguous?

    \item Are the testing objectives consistent with \progname{}'s intended
    purpose and characteristics?

    \item Are the objectives achievable?

    \item Is it clear how each plan relates to \progname{}'s Software
    Development Life Cycle (SDLC)?

    \item Does the relevant documentation list all of \progname{}'s
    documentation-based SDAs?

    \item Does the relevant documentation list all standards used to inform the
    MTP?

\end{itemize}

\paragraph{Plan Overview}
\begin{itemize}

    \item Is the list of team members complete?

    \item Are team member descriptions accurate with respect to their
    professional skills and proficiencies?

    \item Are the high-level goals assigned to each team member unambiguous?

    \item Are the high-level goals assigned to each team member consistent with
    their professional skills and proficiencies?

    \item Are the high-level goals assigned to each team member reasonable for
    each software development stage?

    \item Is there a plan for each of \progname{}'s SDAs?

    \item Are the verification methods reasonable for each SDA?

    \item Are the roles assigned to team members reasonable with respect to
    their professional skills, proficiencies, and schedule?

    \item Are the SDAs listed as necessary inputs for each plan unambiguous?

    \item Are the inputs to each verification method sufficient to complete it?

    \item Are the expected outcomes of each plan unambiguous?

    \item Are the outputs of each verification plan reasonable expectations?

    \item Is the expected completion time for each plan reasonable?

\end{itemize}

\paragraph{Readability}
\begin{itemize}

    \item Is the document well structured?

    \item Is the writing clear and cohesive?

    \item Is the writing concise?

    \item Do all abbreviations appear in ``Symbols, Abbreviations and
    Acronyms'' (Section~\ref{sec:refs})?

    \item Is the document free of spelling and grammar errors?

\end{itemize}