\subsection{System, Integration, and Unit Test Plan Inspection
Guide}\label{appendix:siutpInspection}
This list is a guide for reviewers in the peer review of the System,
Integration, and Unit Test Plan (SIUTP) document (Section~\ref{sec:mtpVV}). The
list organizes prompts by SIUTP section in the order that they appear in the
document. Reviewers are encouraged to include their own checks to this guide.

\paragraph{Introduction}
\begin{itemize}

    \item Does the introduction describe the purpose of \progname{}'s SIUTP?

    \item Does the summary of \progname{} match that of Master Test Plan (MTP)?

    \item Are the aspects of \progname{} that are to be tested by the SIUTP
    unambiguous?

    \item Are the aspects of \progname{} that are \textit{not} to be tested by
    the SIUTP unambiguous?

    \item Are the testing objectives consistent with the description in the MTP?

    \item Are the objectives achievable?

    \item Is it clear how the SIUTP impacts \progname{}'s development?

    \item Does the relevant documentation list all \progname{} documentation
    relevant to the SIUTP?

    \item Does the relevant documentation list all standards used to inform the
    SIUTP?

    \item Is the list of tools for automated testing and verification complete?

    \item Is the list of tools for automated testing and verification
    unambiguous?

\end{itemize}

\paragraph{System Test Description}
\begin{itemize}

    \item Are the inputs to each test case specific?

    \item Are the expected outputs of each test case explicitly defined?

    \item Is the purpose of each test case unambiguous?

    \item Are the methods for quantifying errors unambiguous?

    \item Are the methods for measuring qualitative data unambiguous?

    \item Are the test cases traceable to \progname{}'s requirements as
    described in the Software Requirements Specification (SRS)?

    \item Do the test cases represent a complete coverage of \progname{}'s
    requirements as described in the SRS? If not, does it explain why?

\end{itemize}

\paragraph{Integration Test Description}
\begin{itemize}

    \item Are the inputs to each test case specific?

    \item Are the expected outputs of each test case explicitly defined?

    \item Is the purpose of each test case unambiguous?

    \item Are the methods for quantifying errors unambiguous?

    \item Are the methods for measuring qualitative data unambiguous?

    \item Are the test cases traceable to the relationships between
    \progname{}'s modules as described in the the Module Guide (MG)?

    \item Do the test cases represent a complete coverage of the relationships
    between \progname{}'s modules as described in MG? If not, does it explain
    why?

\end{itemize}

\paragraph{Unit Test Description}
\begin{itemize}

    \item Are the inputs to each test case specific?

    \item Are the expected outputs of each test case explicitly defined?

    \item Is the purpose of each test case unambiguous?

    \item Are the methods for quantifying errors unambiguous?

    \item Are the methods for measuring qualitative data unambiguous?

    \item Are the test cases traceable to the constants, types, and/or access
    routines in \progname{}'s modules as described in the the Module Interface
    Specification (MIS)?

    \item Do the test cases represent a complete coverage of the constants,
    types, and/or access routines in \progname{}'s modules as described in the
    the MIS? If not, does it explain why?

\end{itemize}

\paragraph{Readability}
\begin{itemize}

    \item Is the document well structured?

    \item Is the writing clear and cohesive?

    \item Is the writing concise?

    \item Do all symbols, abbreviations, and acronyms in the SIUTP appear in
    the Reference Material?

    \item Is the document free of spelling and grammar errors?

\end{itemize}