\subsection{SRS Inspection Guide}\label{appendix:srsInspection}
This list is a guide for reviewers in the peer review of the Software
Requirements Specification (SRS) document (Section~\ref{sec:srsVV}). The list
organizes prompts by SRS section in the order that they appear in the document.
Reviewers are encouraged to include their own checks to this guide.

\paragraph{Introduction}
\begin{itemize}

    \item Does the introduction describe the problem domain?

    \item Is the list of stakeholders complete?

    \item Is the description for each stakeholder accurate?

    \item Are the interests of each stakeholder accurate?

    \item Is the document's purpose described unambiguously?

    \item Is the document's purpose described accurately?

    \item Are the characteristics of the intended reader describing the readers
    of the SRS?

    \item Are the characteristics of the intended reader complete?

    \item Is the document accessible to its intended readers?

    \item Does the scope clearly state what \progname{} will do?

    \item Does the scope state clearly what \progname{} will not do?

    \item Is the description of the SRS's document structure accurate?

\end{itemize}

\paragraph{General System Description}
\begin{itemize}

    \item Is the list of entities in the system context clear?

    \item Is the list of entities in the system context complete?

    \item Are the responsibilities for each entity in the system context clear?

    \item Are the responsibilities for each entity in the system context
    complete?

    \item Is the system context diagram accurate?

    \item Are the user characteristics unambiguous?

    \item Are the user characteristics specific?

    \item Are the system constraints accurate?

    \item Are the system constraints complete?

\end{itemize}

\paragraph{Specific System Description: Problem Description}
\begin{itemize}

    \item Is the terminology and definitions list complete?

    \item Is the physical system description complete?

    \item Is the physical system description accurate?

    \item Are the goal statements abstract?

    \item Are the goal statements a refinement of the ``Scope of the
    Requirements'' section and/or stakeholder needs?

    \item Are the goal statements understandable by non-experts of the domain?

\end{itemize}

\paragraph{Specific System Description: Solution Characteristics Description}
\begin{itemize}

    \item Is the choice of affective theories/models traceable to the goal
    statements?

    \item Does each assumption address a single concern?

    \item Are the assumptions traceable to the ``Scope of the Requirements''
    section and/or stakeholder needs?

    \item Is each assumption referenced at least once?

    \item Is the purpose of the Conceptual Models clear?

    \item Are the Theoretical Models refinements of the Conceptual Models?

    \item Is the list of General Definitions complete?

    \item Are derivations of General Definitions clear?

    \item Is the list of Data Definitions complete?

    \item Are derivations of Data Definitions clear?

    \item Are the Type Definitions abstract?

    \item Are the Type Definitions unambiguous?

    \item Are the Type Definitions referenced by at least one other model/type
    definition?

    \item Is it clear which Type Definitions users are responsible for defining?

    \item Are the Instance Models refinements of other models?

    \item Are the Instance Models abstract?

    \item Are the Instance Models clear?

    \item Are there links between each model/type definition and the
    models/type definitions they are referenced by?

    \item Are there links between each model/type definition and the
    models/type definitions they depend on?

    \item Are the Data Constraints complete?

    \item Are the Data Constraints unambiguous?

    \item Are the Properties of a Correct Solution complete?

    \item Are the Properties of a Correct Solution unambiguous?

\end{itemize}

\paragraph{Requirements}
\begin{itemize}

    \item Are the functional requirements abstract?

    \item Are the functional requirements testable?

    \item Are all functional requirements traceable to a stakeholder
    requirement?

    \item Do the nonfunctional requirements have objective fit criteria?

    \item Are the nonfunctional requirement fit criteria testable?

    \item Are all nonfunctional requirements traceable to a stakeholder
    requirement and/or user characteristic?

\end{itemize}

\paragraph{Future Changes}
\begin{itemize}

    \item Are the likely changes traceable to some component of the SRS?

    \item Is it feasible to hide the likely changes in a design?

\end{itemize}

\paragraph{Traceability Matrices and Graphs}
\begin{itemize}

    \item Are all necessary traceability matrices and graphs present?

    \item Are all necessary traceability matrices and graphs accurate?

\end{itemize}

\paragraph{Readability}
\begin{itemize}

    \item Does the document describe the problem domain?

    \item Is the document well structured?

    \item Is the writing clear and cohesive?

    \item Is the writing concise?

    \item Do all symbols, abbreviations, and notation in the SRS appear in the
    Reference Material?

    \item Is the document free of spelling and grammar errors?

\end{itemize}