\subsection{Acceptance Test Plan Inspection
Guide}\label{appendix:validationInspection}
This list is a guide for reviewers in the peer review of the Acceptance Test
Plan (ATP) document (Section~\ref{sec:mtpVV}). The list organizes prompts by
ATP section in the order that they appear in the document. Reviewers are
encouraged to include their own checks to this guide.

\paragraph{Introduction}
\begin{itemize}

    \item Does the introduction describe the purpose of \progname{}'s ATP?

    \item Does the summary of \progname{} match that of Master Test Plan (MTP)?

    \item Are the aspects of \progname{} that are to be tested by the ATP
    unambiguous?

    \item Are the aspects of \progname{} that are \textit{not} to be tested by
    the ATP unambiguous?

    \item Are the testing objectives consistent with the description in the MTP?

    \item Are the objectives achievable?

    \item Is it clear how the ATP impacts \progname{}'s development?

    \item Does the relevant documentation list all \progname{} documentation
    relevant to the ATP?

    \item Does the relevant documentation list all standards used to inform the
    ATP?

\end{itemize}

\paragraph{Test Case Descriptions}
\begin{itemize}

    \item Is the purpose of the test cases unambiguous?

    \item Is the purpose of the test cases traceable to \progname{}'s concept
    summary?

    \item Is the purpose of the test cases traceable to \progname{}'s users and
    user needs?

    \item Are the inputs to each test case specific?

    \item Are the expected outputs of each test case explicitly defined?

    \item Is the purpose of each test case unambiguous?

    \item Are the methods for quantifying errors unambiguous?

    \item Are the methods for measuring qualitative data unambiguous?

    \item Are the test cases traceable to \progname{}'s concept summary?

    \item Are the test cases traceable to \progname{}'s users and user needs?

    \item Do the number and type of test cases seem sufficient?

\end{itemize}

\paragraph{User Study Descriptions}
\begin{itemize}

    \item Is the purpose of the user study unambiguous?

    \item Is the purpose of the user study traceable to \progname{}'s concept
    summary?

    \item Is the purpose of the user study traceable to \progname{}'s users and
    user needs?

    \item Are descriptions of the anticipated number and type of study
    participants unambiguous?

    \item Are the descriptions of study apparatuses complete?

    \item Are the descriptions of study apparatuses unambiguous?

    \item Are the descriptions of study procedures complete?

    \item Are the descriptions of study procedures unambiguous?

    \item Are the descriptions of independent variables complete?

    \item Are the descriptions of independent variables unambiguous?

    \item Are the descriptions of dependent variables complete?

    \item Are the descriptions of dependent variables unambiguous?

    \item Are descriptions of anticipated control, random, and confounding
    variables unambiguous?

    \item Is the user study replicable?

    \item Do the number and type of user studies seem sufficient?

\end{itemize}