\subsection{Implementation Inspection
Guide}\label{appendix:verificationInspection}
This list is a guide for reviewers in the peer review of the source code and
source code documentation (Section~\ref{plan:implementation}). The list
includes prompts for reviewing test case documentation, source code, and source
code documentation. Reviewers are encouraged to include their own checks to
this guide.

\paragraph{Test Case Documentation}
\begin{itemize}

    \item Is the documentation's purpose described unambiguously?

    \item Is the documentation's purpose described accurately?

    \item Is the scope of the test case verification accurate?

    \item Are the types of test cases used accurate?

    \item Are the types of test cases used sufficient?

    \item Are automated testing and/or verification tools documented?

    \item Do the test cases have specific inputs?

    \item Do the test cases have explicitly defined expected outputs?

    \item Is every functional requirement from the SRS satisfied by at least
    one system-level test case?

    \item Is there at least one test case where a set of modules is tested as
    an integrated unit such that only one new module is added to the set in a
    single test case?

    \item Are there sufficient test cases for each module?

\end{itemize}

\paragraph{Source Code and Documentation}
\begin{itemize}

    \item Does each source code file have a header listing the:
    \begin{itemize}
        \item Last modified date
        \item Programming language and its version
        \item Imported libraries and/or components and their version
        \item Operating System and its version
        \item License
        \item Author(s)
    \end{itemize}

    \item Does each function have a header describing its purpose, inputs, and
    outputs?

    \item Are the coding standards of the implementation language used
    consistently throughout the source code and its documentation?

    \item Are constants named consistently?

    \item Are constants given symbolic names?

    \item Are variables named consistently?

    \item Are functions named consistently?

    \item Are source code files organized consistently with respect to each
    other?

    \item Are source code files formatted consistently with respect to each
    other?

    \item Is the source code as simple as possible?

    \item Are modifications to the source code clearly documented?

    \item Is the source code documentation complete/traceable to the source
    code?

    \item Is the source code documentation unambiguous?

\end{itemize}

\paragraph{Readability}
\begin{itemize}

    \item Is the code/document well structured?

    \item Is the writing clear and cohesive?

    \item Is the writing concise?

    \item (If applicable) Do all symbols, abbreviations, and notation in the
    document appear/have a reference to a definition in the Reference Material?

    \item Is the code/document free of spelling and grammar errors?

\end{itemize}